---
title: "R21 Prelim ProjectR Analysis"
date: "210302"
output: html_document

---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(projectR)
library(CoGAPS)
library(ggplot2)
library(DESeq2)
library(gplots)
library(dplyr)
library(tibble)
library(AnnotationDbi)
library(devtools)
library(ggbiplot)
library(biomaRt)


```



# START 
    Goal/Purpose of operations: 
    Publicly available mouse and human datasets for exercise/nonexercise, used to project mouse onto human data using ProjectR
    
    Finished psedocode on: Date
    
    System which operations were done on: MacBook Pro (16-inch, 2019),  Processor 2.4 GHz 8-Core Intel Core i9, Memory 64 GB 2667 MHz DDR4, Graphics AMD Radeon Pro 5300M 4 GB Intel UHD Graphics 630 1536 MB
    
    GitHub Repo:
    
    Directory of operations: /path
    
    Scripts being edited for operations: filename(s)
    
    Data being used: description and location
    
    Papers and tools: deseq with paper 

# STEPS
### Set working directory 
### load in data
### Analysis
comments, comments

### Save Data
### Save Figures
    
# END
    Location of final scripts:
    
    Location of data produced:
    
    Dates when operations were done:
    
## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r, copy = TRUE}
#Mouse data read-in
mouse_data <- read.delim("mouse_fixed_raw_counts.txt", header = TRUE, row.names = 1, sep = "\t", stringsAsFactors = FALSE)
mouse_data <- mouse_data[, -c(1:3)]
mouse_info <- read.csv("mouse_info.csv", header = TRUE)
mouse_metadata <- read.delim("SraRunTable.txt", header = TRUE, sep = ",")
mouse_meta <- mouse_metadata[, c(1, 14, 16, 18)]
#replace white space with underscores in "diet" column
mouse_meta$diet <- gsub(" ", "_", mouse_meta$diet)
```

```{r}
#Normalization
#t_data <- mouse_data %>% t() %>% left_join(mouse_meta, ., by = "Run")
t_data <- t(mouse_data)
t_data <- t_data %>% as.data.frame() %>% rownames_to_column(., var = "Run")
t_data <- left_join(mouse_meta, t_data, by = "Run")
t_data$diet <- gsub(" ", "_", t_data$diet)
mouse_dds <- DESeqDataSetFromMatrix(mouse_data, colData = mouse_meta, design = ~ Group)
mouse_dds <- DESeq(mouse_dds)
mouse_res <- results(mouse_dds)
mouse_vsd <- vst(mouse_dds, blind = FALSE)
mouse_deseq2_pca <- DESeq2::plotPCA(mouse_vsd, intgroup = "diet")
mouse_deseq2_pca
mouse_vsd_matrix <- assay(mouse_vsd)
rownames(mouse_vsd_matrix) <- sub("\\..*", "", rownames(mouse_vsd_matrix))
#back to df and make ensembl gene rownames into a column
mouse_vsd2 <- mouse_vsd_matrix %>% as.data.frame() %>% rownames_to_column(., "ENSEMBL")
```

```{r, copy = TRUE}
#annotation
library(org.Mm.eg.db)
#retrieve conversion info from one ID type to another
symbols_mouse <- AnnotationDbi::select(org.Mm.eg.db, keys = rownames(mouse_vsd_matrix), columns = c("SYMBOL"),keytype = "ENSEMBL")
#determine indices for non-NA genes
non_na_symbols <- which(is.na(symbols_mouse$SYMBOL) == FALSE)
#return only the genes with annotations using indices
symbols_mouse <- symbols_mouse[non_na_symbols, ]
#determine indices for non-duplicated genes
no_dups_symbols <- which(duplicated(symbols_mouse$SYMBOL) == FALSE)
#return only non-dup genes using indices
symbols_mouse <- symbols_mouse[no_dups_symbols, ]
mouse_symbol_data <- inner_join(symbols_mouse, mouse_vsd, by = "ENSEMBL")
mouse_symbol_data <- column_to_rownames(mouse_symbol_data, var = "SYMBOL")
mouse_symbol_data <- mouse_symbol_data[,-1]
```

```{r}
# Basic function to convert mouse to human gene names
convertMouseGeneList <- function(x){
require("biomaRt")
human = useMart("ensembl", dataset = "hsapiens_gene_ensembl")
mouse = useMart("ensembl", dataset = "mmusculus_gene_ensembl")
genesV2 = getLDS(attributes = c("mgi_symbol"), filters = "mgi_symbol", values = x , mart = mouse, attributesL = c("hgnc_symbol"), martL = human, uniqueRows=T)
return(genesV2)
}
```

```{r}
#Use convertMouseGeneList to convert to human genes
mouse_to_human_genes <- convertMouseGeneList(symbols_mouse$SYMBOL)
converted_mouse_data <- mouse_symbol_data %>% rownames_to_column(., var = "MGI.symbol") %>% left_join(., mouse_to_human_genes, by = "MGI.symbol")
#determine non-NA genes
non_na <- which(is.na(converted_mouse_data$HGNC.symbol) == FALSE)
#return only the genes with annotations using indices
converted_mouse_data <- converted_mouse_data[non_na, ]
#determine indices for non-duplicated genes
no_dups <- which(duplicated(converted_mouse_data$HGNC.symbol) == FALSE)
#return only non-dup genes using indices
converted_mouse_data <- converted_mouse_data[no_dups, ]
rownames(converted_mouse_data) <- NULL 
converted_mouse_data <- converted_mouse_data %>% as.data.frame() %>% column_to_rownames(., var = "HGNC.symbol") 
converted_mouse_data <- converted_mouse_data[,c(3:20)]
```


Read in human data, normalize, annotate, look at PCA

```{r}
#HUMAN data read in
human_data <- read.delim("human_fixed_raw_counts.txt", header = TRUE, row.names = 1, sep = "\t")
human_data <- human_data[, -c(1:3)]
human_info <- read.csv("human_info.csv", header = TRUE, sep = ",")
human_info <- human_info[c(1:58),]

#HUMAN normalization
human_dds <- DESeqDataSetFromMatrix(human_data, colData = human_info, design = ~ BMI)
human_dds <- DESeq(human_dds)
human_res <- results(human_dds)
human_vsd <- vst(human_dds, blind = FALSE)
human_vsd_matrix <- assay(human_vsd)
human_deseq2_pca <- plotPCA(human_vsd, intgroup = "BMI")
human_deseq2_pca
rownames(human_vsd_matrix) <- sub("\\..*", "", rownames(human_vsd_matrix))
#back to df and make ensembl gene rownames into a column
human_vsd_matrix <- human_vsd_matrix %>% as.data.frame() %>% rownames_to_column(., "ENSEMBL")

#annotation
library(org.Hs.eg.db)
#retrieve conversion info from one ID type to another
symbols_human <- AnnotationDbi::select(org.Hs.eg.db, keys = human_vsd_matrix$ENSEMBL, columns = c("SYMBOL"),keytype = "ENSEMBL")
#determine indices for non-NA genes
non_na <- which(is.na(symbols_human$SYMBOL) == FALSE)
#return only the genes with annotations using indices
symbols_human <- symbols_human[non_na, ]
#determine indices for non-duplicated genes
no_dups_human <- which(duplicated(symbols_human$SYMBOL) == FALSE)
#return only non-dup genes using indices
symbols_human <- symbols_human[no_dups_human, ]
human_symbol_data <- inner_join(symbols_human, human_vsd_matrix, by = "ENSEMBL")
human_symbol_data <- column_to_rownames(human_symbol_data, var = "SYMBOL")
human_symbol_data <- human_symbol_data[,-1]



```
PCA for mouse data

```{r}
#MOUSE PCA
pc_mouse <- prcomp(t(mouse_vsd_matrix))
#find variance
pc_var <- round(((pc_mouse$sdev)^2/sum(pc_mouse$sdev^2))*100,2)
mouse_pca_df <- data.frame(cbind(pc_mouse$x, mouse_meta))

pc_genotype <-ggplot(mouse_pca_df, aes(x=PC1,y=PC2,color= Genotype ))
pc_genotype <-pc_genotype+geom_point()
pc_genotype
ggsave("Mouse_PC_Genotype.pdf", pc_genotype)

pc3pc4_group <-ggplot(mouse_pca_df, aes(x=PC7,y=PC8,color= Group ))
pc3pc4_group <-pc3pc4_group+geom_point()
pc3pc4_group
ggsave("Mouse_PCA3_4_Group.pdf", pc3pc4_group)

library(factoextra)
#PCA of individual mice
mouse_ind_pca <- fviz_pca_ind(pc_mouse, repel = TRUE)
mouse_ind_pca



```

NMF method for ProjectR

```{r}

#MOUSE pNMF

#the output varies greatly. trying to change parameters to make a better model
params <- new("CogapsParams")
params <- setParam(params, "seed", 1000)

#CoGAPS to find patterns in the data
AP.mouse_seed1000 <- CoGAPS(mouse_symbol_data, params, nIterations = 1000)
#heatmap
pNMF <-heatmap.2(as.matrix(AP.mouse_seed1000),col=bluered, trace='none',
                 distfun=function(c) as.dist(1-cor(t(c))) ,
                 cexCol=1,cexRow=.6,scale = "row", main = "seed 1000, nIterations 1500", labRow = rownames(mouse_symbol_data), xlab = "Patterns",
                 hclustfun=function(x) hclust(x, method="average"))
plot(AP.mouse_seed800)
pNMF



AP.human <- CoGAPS(human_symbol_data, params, nIterations = 1500)
#heatmap
human_NMF <-heatmap.2(as.matrix(AP.human),col=bluered, trace='none',
                 distfun=function(c) as.dist(1-cor(t(c))) ,
                 cexCol=1,cexRow=.6,scale = "row", main = "Patterns in Exercise/Non in Humans", labRow = rownames(human_symbol_data), xlab = "Patterns",
                 hclustfun=function(x) hclust(x, method="average"))
plot(AP.human)



#Data to project into PC's from mouse expression data (Projecting mouse onto human data? ) 
human_data <- as.matrix(human_symbol_data)
human_NMF <- projectR(human_data,loadings=AP.mouse_seed1000, full=TRUE,
                      dataNames=rownames(human_data))
#[1] "30 row names matched between data and loadings"
#[1] "Updated dimension of data: 30 58"
human_meta <- column_to_rownames(human_info, var = "SRR")
dNMF <- human_NMF %>% data.frame() %>% t() %>% cbind(., human_meta)

projection_NMF <- human_NMF$projection %>% t() %>% cbind(., human_meta)
pval_NMF <- human_NMF$pval
#plot pca
projection_plot <- ggplot(projection_NMF, aes(x = Pattern_1, y = Pattern_2, colour = PrePost)) + geom_point() + ggtitle("NMF Mouse Patterns Projected onto Human Data")
projection_plot
ggsave("NMF_Mouse_Projected_to_Human.pdf", projection_plot)

```
PCA Method of Projecting

```{r}
human_deseq2_pc3 <- prcomp(t(human_symbol_data))
pc_mouse_conv <- prcomp(t(converted_mouse_data))
#find variance
pc_var_conv <- round(((pc_mouse_conv$sdev)^2/sum(pc_mouse_conv$sdev^2))*100,2)
mouse_pca_df_conv <- data.frame(cbind(pc_mouse_conv$x, mouse_meta))

PCA_projectr <- projectR(human_data, loadings = pc_mouse_conv, full = TRUE, dataNames = rownames(human_data))
human_meta_t <- t(human_meta)
PCA_projectr_t <- t(PCA_projectr[[1]])
PCA_projectr_df <- cbind(PCA_projectr_t, human_meta)
#[1] "18282 row names matched between data and loadings"
#[1] "Updated dimension of data: 18282 58"

#Plot PCA
dPCA <- data.frame(cbind(t(PCA_projectr[[1]]),PCA_projectr_df))
projected_PCA <- ggplot(PCA_projectr_df, aes(x = PC1, y = PC2, colour = PrePost)) + geom_point() + ggtitle("PCA Mouse Patterns Projected onto Human Data")
projected_PCA
ggsave("PCA_Method_Mouse_Projected_onto_Human.pdf", projected_PCA)

```

```{r}
#cluster methods
mouse_dist <- dist(converted_mouse_data, method = "euclidean")
mouse_hclust <- hclust(mouse_dist, method = "average")

mouse_clusters <- cluster2pattern(clusters = mouse_hclust, NP = 7, mouse_dist)




```

# NEW mouse dataset
```{r, copy = TRUE}
##  TEST 
test_mouse_data <- read.delim("mouse210321_fixed_raw_counts.txt", header = TRUE, row.names = 1, sep = "\t", stringsAsFactors = FALSE)
test_mouse_data <- test_mouse_data[, -c(1:3)]
#metadata not included with this dataset.. 
test_info <- read.csv("MouseSraRunTable.txt", header = TRUE, sep = ",")
Name <- c("CON_pre4","HFD_pre2","HFD_pre3","HFD_pre4","CON_post1","CON_post2","CON_post3","CON_post4","HFD_post4","CON_pre3","CON_pre2","HFD_post2","HFD_post3","HFD_pre1","CON_pre1","HFD_post1")
test_info <- cbind(test_info, Name)
test_info <- mutate(test_info, Diet = ifelse(grepl("HFD", Name, ignore.case = TRUE), "HFD", "CON"))
test_info <- mutate(test_info, Exercise = ifelse(grepl("pre", Name, ignore.case = TRUE), "pre", "post"))
test_dds <- DESeqDataSetFromMatrix(test_mouse_data, colData = test_info, design = ~ Diet + Exercise)
test_dds <- DESeq(test_dds)
test_res <- results(test_dds)
test_vsd <- vst(test_dds, blind = FALSE)
test_deseq2_pca <- DESeq2::plotPCA(test_vsd, intgroup = "Exercise")
test_deseq2_pca
test_vsd_matrix <- assay(test_vsd)
rownames(test_vsd_matrix) <- sub("\\..*", "", rownames(test_vsd_matrix))
test_vsd2 <- test_vsd_matrix %>% as.data.frame() %>% rownames_to_column(., "ENSEMBL")
write.csv(test_info, file= "New_Mouse_Metadata")

#annotation
library(org.Mm.eg.db)
#retrieve conversion info from one ID type to another
test_mouse <- AnnotationDbi::select(org.Mm.eg.db, keys = rownames(test_vsd_matrix), columns = c("SYMBOL"),keytype = "ENSEMBL")
#determine indices for non-NA genes
testnon_na_symbols <- which(is.na(test_mouse$SYMBOL) == FALSE)
#return only the genes with annotations using indices
test_mouse <- test_mouse[testnon_na_symbols, ]
#determine indices for non-duplicated genes
testno_dups_symbols <- which(duplicated(test_mouse$SYMBOL) == FALSE)
#return only non-dup genes using indices
test_mouse <- test_mouse[testno_dups_symbols, ]
test_symbol_data <- inner_join(test_mouse, test_vsd2, by = "ENSEMBL")
test_symbol_data <- column_to_rownames(test_symbol_data, var = "SYMBOL")
test_symbol_data <- test_symbol_data[,-1]

## Basic function to convert mouse to human gene names
convertMouseGeneList <- function(x){
require("biomaRt")
human = useMart("ensembl", dataset = "hsapiens_gene_ensembl")
mouse = useMart("ensembl", dataset = "mmusculus_gene_ensembl")
genesV2 = getLDS(attributes = c("mgi_symbol"), filters = "mgi_symbol", values = x , mart = mouse, attributesL = c("hgnc_symbol"), martL = human, uniqueRows=T)
return(genesV2)
}

#Use convertMouseGeneList to convert to human genes
mouse_to_human_genes_new <- convertMouseGeneList(test_mouse$SYMBOL)
conv_new_mouse_data <- test_symbol_data %>% rownames_to_column(., var = "MGI.symbol") %>% left_join(., mouse_to_human_genes_new, by = "MGI.symbol")
#determine non-NA genes
non_na_new <- which(is.na(conv_new_mouse_data$HGNC.symbol) == FALSE)
#return only the genes with annotations using indices
conv_new_mouse_data <- conv_new_mouse_data[non_na_new, ]
#determine indices for non-duplicated genes
no_dups_new <- which(duplicated(conv_new_mouse_data$HGNC.symbol) == FALSE)
#return only non-dup genes using indices
conv_new_mouse_data <- conv_new_mouse_data[no_dups_new, ]
rownames(conv_new_mouse_data) <- NULL 
conv_new_mouse_data <- conv_new_mouse_data %>% as.data.frame() %>% column_to_rownames(., var = "HGNC.symbol") 
conv_new_mouse_data <- conv_new_mouse_data[,c(2:17)]


#MOUSE NMF
#the output varies greatly. trying to change parameters to make a better model
params <- new("CogapsParams")
params <- setParam(params, "seed", 1000)

#CoGAPS to find patterns in the data
AP_mouse <- CoGAPS(conv_new_mouse_data, params, nIterations = 1500)
#heatmap
NMF_mouse <-heatmap.2(as.matrix(AP_mouse),col=bluered, trace='none',
                 distfun=function(c) as.dist(1-cor(t(c))) ,
                 cexCol=1,cexRow=.6,scale = "row", main = "NEW Mouse Data: seed 1000, nIterations 1500", labRow = rownames(conv_new_mouse_data), xlab = "Patterns",
                 hclustfun=function(x) hclust(x, method="average"))
pdf("NewMouse_NMF_seed1000_nIt1000_1.pdf", NMF_mouse)
plot(AP_mouse, main = "Expression and Sample Weights")
NMF_mouse

#Data to project into PC's from mouse expression data (Projecting mouse onto human data? ) 
human_data <- as.matrix(human_symbol_data)
human_NMF <- projectR(human_data,loadings=AP_mouse, full=TRUE,
                      dataNames=rownames(human_data))
#[1] "18330 row names matched between data and loadings"
#[1] "Updated dimension of data: 18330 58"

human_meta <- column_to_rownames(human_info, var = "SRR")
#dNMF <- human_NMF %>% data.frame() %>% t() %>% cbind(., human_meta)

projection_NMF <- human_NMF$projection %>% t() %>% cbind(., human_meta)
pval_NMF <- human_NMF$pval
#plot pca
projection_plot <- ggplot(projection_NMF, aes(x = Pattern_1, y = Pattern_2, colour = PrePost)) + geom_point() + ggtitle("NEW Mouse Data NMF atterns Projected onto Human Data")
projection_plot
ggsave("NewMouse_NMF_Projected_to_Human_210412.pdf", projection_plot)

#PC Projection Method with New Mouse Data
pc_new_mouse <- prcomp(t(conv_new_mouse_data))
#find variance
pc_var_new_mouse <- round(((pc_new_mouse$sdev)^2/sum(pc_new_mouse$sdev^2))*100,2)
mouse_pca_df_new <- data.frame(cbind(pc_new_mouse$x, test_info))
PCA_projectr_new <- projectR(human_data, loadings = pc_new_mouse, full = TRUE, dataNames = rownames(human_data))
human_meta_t <- t(human_meta)
PCA_projectr_new_t <- t(PCA_projectr_new[[1]])
PCA_projectr_new_df <- cbind(PCA_projectr_new_t, human_meta)
#[1] "30 row names matched between data and loadings"
#[1] "Updated dimension of data: 30 58"

#Plot PCA
dfPCA_new <- data.frame(cbind(t(PCA_projectr_new[[1]]),PCA_projectr_new_df))
projected_PCA_new <- ggplot(PCA_projectr_new_df, aes(x = PC1, y = PC2, colour = PrePost)) + geom_point() + ggtitle("NEW Mouse PC Patterns Projected onto Human Data")
projected_PCA_new
ggsave("New_Mouse_PCA_Method_Mouse_Projected_onto_Human.pdf", projected_PCA_new)




#looking just at PCA of new mouse data
pc_var <- round(((pc_mouse$sdev)^2/sum(pc_mouse$sdev^2))*100,2)
new_mouse_pca_df <- data.frame(cbind(pc_new_mouse$x, test_info))

pc_newmouse_2_3 <- ggplot(new_mouse_pca_df, aes(x=PC3, y=PC4, color= Exercise))
pc_newmouse_2_3 <- pc_genotype+geom_point() + ggtitle("New Mouse Dataset PC2/PC3")
pc_newmouse_2_3
ggsave("New_Mouse_PC_Genotype.pdf", pc_genotype)

```



## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
