---
title: "R21 Prelim ProjectR Analysis"
date: "210302"
output:
  html_document:
    toc: true
    theme: united
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(projectR)
library(CoGAPS)
library(ggplot2)
library(DESeq2)
library(gplots)
library(dplyr)
library(tibble)
library(AnnotationDbi)
library(devtools)
library(ggbiplot)

```



# START 
    Goal/Purpose of operations: 
    Publicly available mouse and human datasets for exercise/nonexercise, used to project mouse onto human data using ProjectR
    
    Finished psedocode on: Date
    
    System which operations were done on: MacBook Pro (16-inch, 2019),  Processor 2.4 GHz 8-Core Intel Core i9, Memory 64 GB 2667 MHz DDR4, Graphics AMD Radeon Pro 5300M 4 GB Intel UHD Graphics 630 1536 MB
    
    GitHub Repo:
    
    Directory of operations: /path
    
    Scripts being edited for operations: filename(s)
    
    Data being used: description and location
    
    Papers and tools: deseq with paper 

# STEPS
### Set working directory 
### load in data
### Analysis
comments, comments

### Save Data
### Save Figures
    
# END
    Location of final scripts:
    
    Location of data produced:
    
    Dates when operations were done:
    
## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r}
#Mouse data read-in
mouse_data <- read.delim("mouse_fixed_raw_counts.txt", header = TRUE, row.names = 1, sep = "\t", stringsAsFactors = FALSE)
mouse_data <- mouse_data[, -c(1:3)]
mouse_info <- read.csv("mouse_info.csv", header = TRUE)
mouse_metadata <- read.delim("SraRunTable.txt", header = TRUE, sep = ",")
mouse_meta <- mouse_metadata[, c(1, 14, 16, 18)]
#replace white space with underscores in "diet" column
mouse_meta$diet <- gsub(" ", "_", mouse_meta$diet)

#Normalization
#t_data <- mouse_data %>% t() %>% left_join(mouse_meta, ., by = "Run")
t_data <- t(mouse_data)
t_data <- t_data %>% as.data.frame() %>% rownames_to_column(., var = "Run")
t_data <- left_join(mouse_meta, t_data, by = "Run")
t_data$diet <- gsub(" ", "_", t_data$diet)
mouse_dds <- DESeqDataSetFromMatrix(mouse_data, colData = mouse_meta, design = ~ Group)
mouse_dds <- DESeq(mouse_dds)
mouse_res <- results(mouse_dds)
mouse_vsd <- vst(mouse_dds, blind = FALSE)
mouse_deseq2_pca <- DESeq2::plotPCA(mouse_vsd, intgroup = "diet")
mouse_deseq2_pca
mouse_vsd_matrix <- assay(mouse_vsd)
rownames(mouse_vsd_matrix) <- sub("\\..*", "", rownames(mouse_vsd_matrix))
#back to df and make ensembl gene rownames into a column
mouse_vsd <- mouse_vsd_matrix %>% as.data.frame() %>% rownames_to_column(., "ENSEMBL")

#annotation
library(org.Mm.eg.db)
#retrieve conversion info from one ID type to another
ont_mouse <- AnnotationDbi::select(org.Mm.eg.db, keys = rownames(mouse_vsd_matrix), columns = c("ONTOLOGY"),keytype = "ENSEMBL")
#determine indices for non-NA genes
non_na_ont <- which(is.na(ont_mouse$ONTOLOGY) == FALSE)
#return only the genes with annotations using indices
ont_mouse <- ont_mouse[non_na_ont, ]
#determine indices for non-duplicated genes
no_dups_ont <- which(duplicated(ont_mouse$ONTOLOGY) == FALSE)
#return only non-dup genes using indices
ont_mouse <- ont_mouse[no_dups_ont, ]
mouse_ont_data <- inner_join(ont_mouse, mouse_vsd_matrix, by = "ENSEMBL")
mouse_ont_data <- column_to_rownames(mouse_ont_data, var = "ONTOLOGY")
mouse_ont_data <- mouse_ont_data[,-1]

#HUMAN data read in
human_data <- read.delim("human_fixed_raw_counts.txt", header = TRUE, row.names = 1, sep = "\t")
human_data <- human_data[, -c(1:3)]
human_info <- read.csv("human_info.csv", header = TRUE, sep = ",")
human_info <- human_info[c(1:58),]

#HUMAN normalization
human_dds <- DESeqDataSetFromMatrix(human_data, colData = human_info, design = ~ BMI)
human_dds <- DESeq(human_dds)
human_res <- results(human_dds)
human_vsd <- vst(human_dds, blind = FALSE)
human_vsd_matrix <- assay(human_vsd)
human_deseq2_pca <- plotPCA(human_vsd, intgroup = "BMI")
human_deseq2_pca
rownames(human_vsd_matrix) <- sub("\\..*", "", rownames(human_vsd_matrix))
#back to df and make ensembl gene rownames into a column
human_vsd_matrix <- human_vsd_matrix %>% as.data.frame() %>% rownames_to_column(., "ENSEMBL")

#annotation
library(org.Hs.eg.db)
#retrieve conversion info from one ID type to another
symbols_human <- AnnotationDbi::select(org.Hs.eg.db, keys = human_vsd_matrix$ENSEMBL, columns = c("SYMBOL"),keytype = "ENSEMBL")
#determine indices for non-NA genes
non_na <- which(is.na(symbols_human$SYMBOL) == FALSE)
#return only the genes with annotations using indices
symbols_human <- symbols_human[non_na, ]
#determine indices for non-duplicated genes
no_dups_human <- which(duplicated(symbols_human$SYMBOL) == FALSE)
#return only non-dup genes using indices
symbols_human <- symbols_human[no_dups_human, ]
human_symbol_data <- inner_join(symbols_human, human_vsd_matrix, by = "ENSEMBL")
human_symbol_data <- column_to_rownames(human_symbol_data, var = "SYMBOL")
human_symbol_data <- human_symbol_data[,-1]

#library(groupdata2)
#human_downsampled <- downsample(human_symbol_data, cat_col = rownames(human_symbol_data))

human_deseq2_pc3 <- prcomp(t(human_symbol_data))
human_pc_df <- data.frame(PC1 = human_deseq2_pc3$x[,1], PC2 = human_deseq2_pc3$x[,2])
ggplot(data = human_pc_df, aes_string(x = "PC1", y = "PC2"))

```
PCA for mouse data

```{r}
#MOUSE PCA
pc_mouse <- prcomp(t(mouse_vsd_matrix))
#find variance
pc_var <- round(((pc_mouse$sdev)^2/sum(pc_mouse$sdev^2))*100,2)
mouse_pca_df <- data.frame(cbind(pc_mouse$x, mouse_meta))

pc_genotype <-ggplot(mouse_pca_df, aes(x=PC1,y=PC2,color= Genotype ))
pc_genotype <-pc_genotype+geom_point()
pc_genotype
ggsave("Mouse_PC_Genotype.pdf", pc_genotype)

pc3pc4_group <-ggplot(mouse_pca_df, aes(x=PC3,y=PC4,color= Group ))
pc3pc4_group <-pc3pc4_group+geom_point()
pc3pc4_group
ggsave("Mouse_PCA3_4_Group.pdf", pc3pc4_group)

library(factoextra)
#PCA of individual mice
mouse_ind_pca <- fviz_pca_ind(pc_mouse, repel = TRUE)
mouse_ind_pca



```

NMF method for ProjectR

```{r}

#MOUSE pNMF

#the output varies greatly. trying to change parameters to make a better model
params <- new("CogapsParams")
params <- setParam(params, "seed", 1000)

#CoGAPS to find patterns in the data
AP.mouse_seed1000 <- CoGAPS(mouse_symbol_data, params, nIterations = 1500)
#heatmap
pNMF <-heatmap.2(as.matrix(AP.mouse_seed1000),col=bluered, trace='none',
                 distfun=function(c) as.dist(1-cor(t(c))) ,
                 cexCol=1,cexRow=.6,scale = "row", main = "seed 1000, nIterations 1500", labRow = rownames(mouse_symbol_data), xlab = "Patterns",
                 hclustfun=function(x) hclust(x, method="average"))
plot(AP.mouse_seed800)
pNMF



AP.human <- CoGAPS(human_symbol_data, params, nIterations = 1500)
#heatmap
human_NMF <-heatmap.2(as.matrix(AP.human),col=bluered, trace='none',
                 distfun=function(c) as.dist(1-cor(t(c))) ,
                 cexCol=1,cexRow=.6,scale = "row", main = "Patterns in Exercise/Non in Humans", labRow = rownames(human_symbol_data), xlab = "Patterns",
                 hclustfun=function(x) hclust(x, method="average"))
plot(AP.human)



#Data to project into PC's from mouse expression data (Projecting mouse onto human data? ) 
human_data <- as.matrix(human_symbol_data)
human_NMF <- projectR(human_data,loadings=AP.mouse_seed1000, full=TRUE,
                      dataNames=rownames(human_data))
#[1] "30 row names matched between data and loadings"
#[1] "Updated dimension of data: 30 58"
human_meta <- column_to_rownames(human_info, var = "SRR")
dNMF <- human_NMF %>% data.frame() %>% t() %>% cbind(., human_meta)

projection_NMF <- human_NMF$projection %>% t() %>% cbind(., human_meta)
pval_NMF <- human_NMF$pval
#plot pca
projection_plot <- ggplot(projection_NMF, aes(x = Pattern_1, y = Pattern_2, colour = PrePost)) + geom_point() + ggtitle("Mouse Patterns Projected onto Human Data")
projection_plot
ggsave("Mouse_Projected_to_Human.pdf", projection_plot)

```
PCA Method of Projecting

```{r}
human_deseq2_pc3 <- prcomp(t(human_symbol_data))
pc_mouse <- prcomp(t(mouse_vsd_matrix))
#find variance
pc_var <- round(((pc_mouse$sdev)^2/sum(pc_mouse$sdev^2))*100,2)
mouse_pca_df <- data.frame(cbind(pc_mouse$x, mouse_meta))

PCA_projectr <- projectR(human_data, loadings = pc_mouse, full = TRUE, dataNames = rownames(human_data))
PCA_projectr_df <- data.frame(Condition = sapply(colnames(human_data), function(x) unlist(strsplit(x, '_'))[1]), stringsAsFactors = FALSE) 
#[1] "0 row names matched between data and loadings"
#[1] "Updated dimension of data: 0 58"

#just for reference, delete afterwards
human_NMF <- projectR(human_data,loadings=AP.mouse_seed1000, full=TRUE,
                      dataNames=rownames(human_data))
human_meta <- column_to_rownames(human_info, var = "SRR")
dNMF <- human_NMF %>% data.frame() %>% t() %>% cbind(., human_meta)

projection_NMF <- human_NMF$projection %>% t() %>% cbind(., human_meta)
pval_NMF <- human_NMF$pval
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
